---
title: "Reading an xlsb workbook"
author: "Michael Allen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading an xlsb workbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(readxlsb)
```

`read_xlsb` attempts to import a region from a binary format Excel workbook (xlsb)

The parameters are 
```{r eval = FALSE}
read_xlsb(path, sheet, range, col_names, col_types, na, trim_ws, skip, ...)
```

## sheet

Either a name, or the index of the sheet to read. Index of the first sheet is 1. If the sheet name is embedded in the range argument, or implied if range is a named range, then this argument is ignored

## range

range can be specified as 

* A named range. Named ranges are not case sensitive
* In **Sheet!A1** notation
* In **Sheet!R1C1** notation
* As a [cellranger::cell_limits](https://cran.r-project.org/web/packages/cellranger/index.html) object

## col_names

* TRUE: The first row is used for column names. Empty cells result in a column name of the form 'column.i' <br />
* FALSE: Column names will be 'column.i' <br />
* Character vector: vector containing column names.

## col_types

Can be implied from the spreadsheet or specified in advanced.
When specifying types, options are

* "*logical*" (or "*boolean*"), "*numeric*" (or "*double*"), "*integer*", "*date*" and "*string*" (or "*character*")
* Use "*skip*" (or "*ignore*") to skip a column
  
When implying types from the underlying spreadsheet data, the resultant type is the regarded as the 'least fragile'. <br /><br />
Effectively the order is *logical* -- *datetime* -- *integer* -- *double* -- *string* <br />

* If 99 rows are of type '*integer*' and 1 row is of type '*double*', then all cells are regarded as '*double*' in that column. <br />
* If 99 rows are of type '*date*' and 1 row is of type '*string*', then all cells are promoted to '*string*' <br />

Currently '*date*' is implied from cell formatting. It is either one of the built-in Excel datetime formats or a custom format where the format string contains only the characters Y, M, D, H, S, y, m, d, h, s and - (dash), : (colon), (space), . (dot). That should be good enough to identify any dates.

If '*date*' is specified as the column type, then any strings are converted to dates. The format I'm afraid isn't flexible at this stage - it's assumed to be of the type "%Y-%m-%dT%H:%M:%S". 

I think there's a 1900 leap year bug in Excel. I haven't corrected for that. 

If any of the cells in a '*date*' column contain time, then a POSIXct object is returned, otherwise a Date object is returned. The timezone for POSIXct is set as UTC. Seems like the sensible thing to do.

## na

A character string that is interpret as NA. This does not effect the implied data type for a column.

## trim_ws

Should leading and trailing whitespaces be trimmed from character strings?

## skip

The number of rows to skip before reading data. 

## ...

Additional options. At present just debug = TRUE is supported. This returns a list with fields '*result*' set to the resulting data.frame and '*env*' set to an internal environment that may be useful for debugging. <br />

```{r eval = TRUE}
res = read_xlsb(path = system.file("extdata", "TestBook.xlsb", package = "readxlsb"), range = "PORTFOLIO", debug = TRUE)

ls(res$env)

res$env$named_ranges
```

